Simulations of Ballot Polling Risk-Limiting Audits

### 2022-04-21 MultipleRoundSims/poster.tex
a

\footercontent{
  \href{odbroadrick@gmail.com}{odbroadrick@gmail.com}
}
% (can be left out to remove footer)

@@

\begin{document}
\addtobeamertemplate{headline}{}
{
    \begin{tikzpicture}[remember picture,overlay]
      \node [anchor=north west, inner sep=3cm] at ([xshift=0.0cm,yshift=3cm]current page.north west)
      {\includegraphics[height=8cm]{logos/gw.png}}; % also try shield-white.eps
      \node [anchor=north east, inner sep=3cm] at ([xshift=0.0cm,yshift=3cm]current page.north east)
      {\includegraphics[height=8cm]{logos/gw.png}};
    \end{tikzpicture}
}

@@

\begin{itemize}
\item
We provide simulations \cite{simulations} of EoR \BRAVO, SO \BRAVO, and \Minerva that help us understand their behavior across multiple rounds and with lower
first round stopping probability; this information can be used to advise election officials.
\end{itemize}

@@

\begin{itemize}
\item 
Software: R2B2 library\cite{r2b2} which has Python implementations of several RLAs as well as a simulator
\item 
Risk limit: 10\%
\item
Margins: 2020 US Presidential election statewide results, limiting ourselves to pairwise margins
for the two main candidates of $0.05$ or larger
\item 
Trials per state: $10,000=10^4$ audits assuming the underlying election is as announced,  
and $10,000=10^4$ audits assuming the underlying election is a tie
\item 
Round schedules: EoR \BRAVO and SO \BRAVO round sizes are chosen to achieve the same
probability of stopping in each round given preceding samples. \Minerva first round sizes
are chosen to achieve some probability of stopping and subsequent round sizes are given by multiplying the previous
round size by $1.5$ or by $1.0$.
\end{itemize}
\end{block}

@@

\begin{figure}[h]
\centering
\begin{minipage}{.49\textwidth}
\includegraphics[width=\textwidth]{prov_risk.png}
\label{fig:bravo_risk}
\end{minipage}
\begin{minipage}{.49\textwidth}
\includegraphics[width=1.0\textwidth]{prov_sprob.png}
\label{fig:minerva_risk}
\end{minipage}
\caption{
The left plot shows the fraction of \Prov audits that stopped in any of the $5$ rounds when the underlying election was a tie
where each round size is selected to achieve $\chi_j=0.9$.
The right plot shows the fraction of \Prov audits that stopped in any of the $5$ rounds when the underlying election was a tie for rounds
$j=1,2,3$ and $\chi_j=0.9$.
}
\label{fig:prov-risk-sprob}
\end{figure}

\begin{figure}
\includegraphics[width=.65\textwidth]{prov_asn.png}
\caption{Cumulative fraction of audits that stopped as a function of average number of sampled ballots for the state of Texas with margin $.057$ with first round stopping probability $\chi_1=0.9$. \Prov, \Minerva with multiplier $1.5$, and both implementations of \BRAVO are included.}
\end{figure}

@@

\separatorcolumn
\end{columns}
\end{frame}
\end{document}

>>>>

\footercontent{
  odbroadrick@gmail.com
}
% (can be left out to remove footer)

@@

\begin{document}
\addtobeamertemplate{headline}{}
{
    \begin{tikzpicture}[remember picture,overlay]
      \node [anchor=north west, inner sep=3cm] at ([xshift=0.0cm,yshift=3cm]current page.north west)
      {\includegraphics[height=9cm]{logos/gw.png}}; % also try shield-white.eps
      \node [anchor=north east, inner sep=3cm] at ([xshift=0.0cm,yshift=3cm]current page.north east)
      {\includegraphics[height=9cm]{logos/gw.png}};
    \end{tikzpicture}
}

@@

\begin{itemize}
\item
We provide simulations~\cite{simulations} of EoR \BRAVO, SO \BRAVO, and \Minerva that help us understand their behavior across multiple rounds and with lower
first round stopping probability; this information can be used to advise election officials.
\end{itemize}

@@

\begin{itemize}
\item 
Software: R2B2 library \cite{r2b2} which has Python implementations of several RLAs as well as a simulator
\item 
Risk limit: 10\%
\item
Margins: 2020 US Presidential election statewide results, limiting ourselves to pairwise margins
for the two main candidates of $0.05$ or larger
\item 
Trials per state: $10,000=10^4$ audits assuming the underlying election is as announced,  
and $10,000=10^4$ audits assuming the underlying election is a tie
\item 
Round schedules: EoR \BRAVO and SO \BRAVO round sizes are chosen to achieve the same
probability of stopping in each round given preceding samples. \Minerva first round sizes
are chosen to achieve some probability of stopping and subsequent round sizes are given by multiplying the previous
round size by $1.5$ or by $1.0$.
\end{itemize}
\end{block}

@@

\begin{figure}[h]
\centering
\begin{minipage}{.49\textwidth}
\includegraphics[width=\textwidth]{prov_risk.png}
\label{fig:prov_risk}
\end{minipage}
\begin{minipage}{.49\textwidth}
\includegraphics[width=1.0\textwidth]{prov_sprob.png}
\label{fig:prov_sprob}
\end{minipage}
\caption{
The left plot shows the fraction of \Prov audits that stopped in any of the $5$ rounds when the underlying election was a tie
where each round size is selected to achieve $\chi_j=0.9$.
The right plot shows the fraction of \Prov audits that stopped in any of the $5$ rounds when the underlying election was a tie for rounds
$j=1,2,3$ and $\chi_j=0.9$.
}
\label{fig:prov-risk-sprob}
\end{figure}

\begin{figure}
\includegraphics[width=.65\textwidth]{prov_asn.png}
\caption{Cumulative fraction of audits that stopped as a function of average number of sampled ballots for the state of Texas with margin $.057$ with first round stopping probability $\chi_1=0.9$. \Prov, \Minerva with multiplier $1.5$, and both implementations of \BRAVO are included.}
\label{fig:prov-asn}
\end{figure}

@@

\end{columns}

\end{frame}
\end{document}

### 2022-04-21 MultipleRoundSims/slides.tex
a

\begin{itemize}
\item Risk-Limiting Audits
\begin{itemize}
\item \BRAVO and \Minerva
\end{itemize}
\item 
\end{itemize}

@@

\begin{frame}
%\begin{abstract}
    In this paper we present simulation results comparing the risk, stopping probability, and number of ballots required over multiple rounds of ballot polling risk-limiting audits (RLAs) \Minerva, Selection-Ordered (SO) \BRAVO, and End-of-Round (EoR) \BRAVO.    
    \BRAVO is the most commonly used ballot polling RLA and requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced. 
    In real audits, multiple ballots are drawn at a time, and \BRAVO is implemented as SO \BRAVO or EoR \BRAVO. 
    \Minerva is a recently proposed ballot polling RLA that requires fewer ballots than either implementation of \BRAVO in a first round with stopping probability $0.9$ but requires a predetermined round schedule.
    %\BRAVO is the most commonly used ballot polling RLA and requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced. In real audits, election officials draw multiple ballots at a time, and \BRAVO rules may be implemented as SO \BRAVO or EoR \BRAVO, with neither being optimal for large first round sizes. \Minerva was recently proposed for use when the first-round stopping probability is large. It requires fewer first-round ballots, on average, than both SO \BRAVO and EoR \BRAVO for first round stopping probabilities of $0.9$ when the (true) underlying election is as announced. 
    It is an open question how these audits compare over multiple rounds and for lower stopping probabilities. 
    Our simulations use stopping probabilities of $0.9$ and $0.25$.
    %In this paper, we present results from simulations of multiple round audits with stopping probabilities of $0.9$ and $0.25$.
    %We examine both: first-round stopping probabilities of $0.9$, a common choice among election officials, and $0.25$, which would be more favorable to \BRAVO. 
    The results are consistent with predictions of the R2B2 open-source library for ballot polling audits. 
    %Because the \Minerva round schedule needs to be predetermined, we observe that the simulator is useful as a tool for understanding the behavior of \Minerva with the fixed round schedule. 
    We observe that both \BRAVO audits are more conservative than \Minerva, which stops with fewer ballots, for both first round stopping probabilities. 
    However, the advantage of using \Minerva decreases considerably for the smaller first round stopping probability, as one would expect.

\keywords{risk-limiting audit (RLA)  \and ballot polling audit \and evidence-based elections \and statistical election audit}
\end{frame}
%\end{abstract}
%
%
\section{Introduction}
\label{sec:intro}
\input{sections/01-intro}

\section{Related work}
\label{sec:related}
\input{sections/02-related}

%\section{Software}
%\label{sec:software}
%\input{sections/03-software}

\section{Experiments}
\label{sec:expts}
\input{sections/04-expts}

\section{Stopping Probability and Risk}
\label{sec:results1}
\input{sections/05-results1}

\section{Number of Ballots}
\label{sec:results2}
\input{sections/06-results2}

\section{Conclusions and Future Work}
\label{sec:conc}
\input{sections/07-conc}

%\section{Acknowledgements}
%\label{sec:ack}
%This research was supported in part by NSF Awards 2015253
%and 1421373.

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{audits}
%
\end{document}

>>>>

\begin{itemize}
\item Risk-Limiting Audits
\begin{itemize}
\item \BRAVO and \Minerva
\end{itemize}
\item Experiments
\item Results
\begin{itemize}
\item Stopping Probability
\item Risk
\item Number of Ballots
\end{itemize}
\item Discussion and Future Work
\end{itemize}

@@

\begin{frame}
\frametitle{Post-tabulation audits}

\begin{itemize}
\item Scanning machines are used to tabulate ballots
\begin{itemize}
\item Cannot trust the machines: bugs, configuration errors, hacking
\end{itemize}
\item Post-tabulation audits
\item Risk-Limiting Audits
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Risk-Limiting Audits}

\begin{itemize}
\item Risk-Limiting Audit (RLA) is a post-tabulation audit that manualy
checks a random sample of voters' ballots
\item Relies on a voter-verified paper trail
\item Sketch:
\begin{enumerate}
\item election results announced
\item sample ballots randomly
\item check if the sample is 'statistically similar' to the announced tally
\begin{itemize}
\item "Yes" - correct - stop the audit
\item "No" - incorrect - proceed to a full hand count
\item "Don't know yet" - undetermined - draw more samples (goto: 2)
\end{itemize}
\end{enumerate}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{\BRAVO}
\begin{itemize}
\item Most commonly used ballot polling RLA
\item In the two candidate case is an instance of Wald's Sequential Probability Ratio Test (SPRT)
\item Is thus the most efficient RLA when ballots are drawn sequentially (i.e. ballot-by-ballot)
\item Real audits are performed in rounds for which \BRAVO can be implemented as:
\begin{itemize}
\item Selection-Ordered (SO) \BRAVO
\item End-of-Round (EoR) \BRAVO
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{\Minerva}
\begin{itemize}
\item Recent RLA designed for round-by-round use
\item Uses a ratio of the \emph{tails} of the probability distributions used in \BRAVO
\item Known to require half the number of ballots as EoR \BRAVO in a first round to achieve a large ($0.90$) probability of stopping
\item Unknown how the audits compare for smaller stopping probability or for rounds after the first
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Experiments}
\begin{itemize}
\item Use simulations to provide evidence for theoretical claims
\item R2B2 software library for round-by-round and ballot-by-ballot RLAs
\item Simulate RLAs for election results from the 2020 Presidential election (all margins above $0.05$)
\begin{itemize}
\item $10000=10^4$ trials assuming the underlying election is as announced
\item $10000=10^4$ trials assuming the underlying election is a tie
\end{itemize}
\item Risk limit: $10\%$
\item Round schedules:
\begin{itemize}
\item \BRAVO round sizes to achieve a chosen probability of stopping in each round given that the audit has already reached that round
\item \Minerva first round sizes to achieve a chosen probability of stopping, and subsequent round sizes found by multiplying the previous round size by a constant ($1.5$ and $1$)
\end{itemize}
\item Stopping probabilities: $0.90$ and $0.25$
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Experiments}
\begin{definition}
An audit $\mathcal{A}$ takes a sample of ballots $X$ as input and gives as output either
(1) $Correct$: the audit is complete, or (2) $Uncertain$: continue the audit.
\end{definition}

\begin{itemize}
\item Binary hypothesis test: $H_0$ (a tie) and $H_a$ (announced results)
\end{itemize}

\begin{definition}[Risk]
The maximum risk $R$ of audit $\mathcal{A}$ with sample $X\in \{0,1\}^*$ drawn from 
the true underlying distribution of ballots is
$R(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct \mid H_0].$
\end{definition}

\begin{definition}[Risk Limiting Audit ($\alpha$-RLA)]
An audit $\mathcal{A}$ is a Risk Limiting Audit with 
risk limit $\alpha$ iff 
$R(\mathcal{A}) \le \alpha.$
\end{definition}
\end{frame}

\begin{frame}
\frametitle{Experiments}
\begin{definition}[Stopping Probability]
The stopping probability $S_j$ of an audit $\mathcal{A}$ in round $j$ is 
$S_j(\mathcal{A})=$
$$\Pr[\mathcal{A}(X)=Correct ~in~round~j~\land \mathcal{A}(X) \neq Correct ~previously \mid H_a]$$
\end{definition}

\begin{definition}[Cumulative Stopping Probability]
The cumulative stopping probability $C_j$ of an audit $\mathcal{A}$ in round $j$ is $C_j(\mathcal{A})= \sum_{i=1}^j S_j$
\end{definition}

\begin{definition}[Conditional Stopping Probability]
The conditional stopping probability  of an audit $\mathcal{A}$ in round $j$ is 
$\chi_j (\mathcal{A})=$
$$\Pr[\mathcal{A}(X)=Correct ~in~round~j~\mid H_a \land \mathcal{A}(X) \neq Correct ~previously]$$
\end{definition}
\end{frame}

\begin{frame}
\frametitle{Results: Stopping Probability ($\chi_1=0.9$)}

\includegraphics[width=0.5\textwidth]{eor_bravo_90perc_10^4_corrected/sprob_first_three_cropped.png}
\includegraphics[width=0.5\textwidth]{so_bravo_90perc_10^4/sprob_first_three.png}

\includegraphics[width=0.5\textwidth]{minerva_multiround_1x_10^4/sprobs_first_three.png}
\includegraphics[width=0.5\textwidth]{minerva_multiround_1p5x_10^4/sprobs_first_three.png}
\end{frame}

\begin{frame}
\frametitle{Results: Stopping Probability ($\chi_1=0.25$)}

\includegraphics[width=1\textwidth]{minerva25percthen1p5_sprob.png}
\end{frame}

\begin{frame}
\frametitle{Results: Risk}

\centering
\noindent \includegraphics[width=1.07\textwidth]{both_risk_plots.png}
\end{frame}

\begin{frame}
\frametitle{Results: Number of Ballots}

\includegraphics[width=0.5\textwidth]{texas25.png}
\includegraphics[width=0.5\textwidth]{missouri25.png}

\includegraphics[width=0.5\textwidth]{massachusetts25.png}
\includegraphics[width=0.5\textwidth]{texas90.png}
\end{frame}

\begin{frame}
\frametitle{Results: Round Size Proportions}

For $\chi_1 = 0.25$, the number of ballots required for \Minerva is smaller than that required by SO \BRAVO and EoR \BRAVO
\begin{itemize}
\item Improvement considerably smaller than that when the stopping probability is $0.9$
\item Number of ballots for SO \BRAVO for $\chi_1 =0.9$ is about a third more than that required by \Minerva, but for $\chi_1 = 0.25$, it requires only about a tenth more ballots than does \Minerva
\item Number of ballots for EoR \BRAVO for $\chi_1 = 0.9$ is about twice those required by \Minerva, but for $\chi_1 =0.25$, it requires only about a fourth to a half more ballots (depending on margin) than does \Minerva

\end{itemize}
For $\chi_1 = 0.9$ with multiplying factor $1$, \Minerva consequent conditional stopping probabilities are about $0.75$ and $0.74$ respectively for rounds two and three. 
\begin{itemize}
\item When the multiplying factor is $1.5$, we see $\chi_2\approx 0.91$
and $\chi_3=0.83$
\end{itemize}

\begin{frame}
\frametitle{Conclusion}

\begin{itemize}

\item We describe use of the R2B2 library and simulator to characterize: 
\begin{itemize}
\item risk, 
\item stopping probability, and 
\item number of ballots 
\item for various round schedules.
\end{itemize}
\item \Minerva requires fewer ballots than either implementation of \BRAVO in all cases we study, but the advantage decreases for a smaller stopping probability

\end{itemize}

\begin{frame}
\frametitle{Future Work}
\begin{itemize}
\item More detailed study of the impact of different round schedules
\item Simulations with other underlying distributions
\end{itemize}
\end{frame}

\begin{frame}
\centering 
{Thank you}
\end{frame}

% References?

\end{document}
